---
title: Twitter Data Analysis with R
author: mutuvi
date: '2017-03-03'
slug: twitter-data-analysis-with-r
categories:
  - '2017'
tags:
  - tidyverse
---

In this blog, I am going to use various R packages mainly from the tidyverse package by [Hadley wickham](http://hadley.nz/) to analyze tweets from twitter archive of  [lordmwesh](https://twitter.com/lordmwesh) to find out his tweeting habits.

[lordmwesh](https://twitter.com/lordmwesh) is an ICT policy expert with vast experience in Internet governance. He was extremely instrumental in my formative stages of my career and continue to walk with me in my journey of data analytics. His twitting habits will be unravelled 

Twitter archive is a nice large dataset with results that can be interesting but not critical to the running of the world. To download your twitter data browse to your Twitter Profile > Settings page, then click **request your archive**. You should be able to download a .zip file with among other files, a nicely formated dataset of your tweets in form of an excel file.

The first step was to load the tweets and quick summary was able to show that lordmwesh had tweeted or retweeted 21447 times from the time he joined twitter

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
library(scales)

tweets <- read.csv("data/tweets.csv", stringsAsFactors = FALSE)
count_tweets <- tweets %>% summarise("Total Tweets"=n())
count_tweets
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
tweets$timestamp <- ymd_hms(tweets$timestamp)
tweets$timestamp <- with_tz(tweets$timestamp, "Africa/Nairobi")

# tweets <- tweets %>% group_by(year(timestamp)) %>% summarize(count=n())
# tweets
  

ggplot(data = tweets, aes(x = year(timestamp))) +
  geom_histogram(breaks = seq(2008.5, 2018.5, by =1), aes(fill = ..count..)) +
  theme(legend.position = "none") +
  xlab("Time") + ylab("Number of tweets") + 
  scale_fill_gradient(low = "midnightblue", high = "aquamarine4")
```

